---
title: "R for Stats"
author: "Alexey Abramov"
date:  "10/9/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

# Setup
```{r, setup}
library(tidyverse)
library(patchwork)
library(hexbin)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(
  ggthemes::theme_fivethirtyeight() + theme(legend.position = "bottom")
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.colour = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Binomial distribution


The probability that a patient fails to follow up in a given month is 15%. 
A pilot study is conducted to assess feasibility involving 5 patients. 

To obtain a binomial probability for a single X value: 
dbinom(x, n, p) 

So for Example 1, we have: dbinom(0, 5, 0.15) 

To obtain a probability for values less than or equal to X:
pbinom(x, n, p) 

So for Example 2, we have: pbinom(2, 5, 0.15)

# Poisson Distribution

Example: 
The emergency room at CUMC receives an average of 7.5 patients each hour. 

To obtain a poisson probability for a single X value in R: 
dpois(x, lambda)

So for (a) above we have: dpois(9, 7.5) 

To obtain a probability for values less than or equal to X in R: 
ppois(x, lambda)

So for (b) above we have: 1 - ppois(3, 7.5) 

# Bootstrapping

## read in the data

```{r}
wloss_data = read_table("./data/Weightloss20.txt")
```

### define the object with the values of interest

```{r}
loss = wloss_data$loss
```

### let's look at a histogram of the loss values
```{r}
hist(loss)
```


### compute the sample statistic from the observed sample
```{r}
xbar <- mean(loss)
xbar
```


### how to draw one bootstrap sample

```{r}
sample(loss, size = 20, replace = T)
```

### Set a seed
you don't get the same sample as above since it is randomly selected to get the same sample, set a seed and run the two lines below together

```{r}
set.seed(1234)
sample(loss, size = 20, replace = T)
```


we want to draw many bootstrap samples, not just one so we need to call the sample function many times; we can sit there and hit "control-return" B times or we can use what's called a "for loop"; let's say B = 50 for now


```{r}
for(j in 1:50) {sample(loss, size = 20, replace = T)}
```

ok it looks like something ran but we don't see anything; that's because we
didn't ask R to store the 50 samples anywhere; so before we use the for loop
we need to create an object with 50 slots to store the 50 samples

```{r}
bs_samps <- vector("list", 50)
```

let's put it together

```{r}
bs_samps <- vector("list", 50)
set.seed(1234)
for(j in 1:50) 
  {bs_samps[[j]] = sample(loss, size = 20, replace = T)}
```

bs_samps contains the 50 bootstrap samples; we now want to compute the 
bootstrap replications - compute the statistic (the mean in this example)  of interest on each bootstrap sample

```{r}
bs_reps = lapply(bs_samps, mean)
```

but it's actually easier to have these values as a vector rather than as a list so we can "unlist" the object

```{r}
bs_reps <- unlist(lapply(bs_samps, mean))
```

now we can analyze the bootstrap replications to get an understanding 
of the sampling distribution of the statistic we are interested in 
(i.e., the sample mean of the weight loss values)

observe the shape of the distribution and standard deviation of the bootstrap replications
```{r}
hist(bs_reps) 
sd(bs_reps) 
```


### Bootstrapping code

first read in the data set that you want to use

X <- "variable of interest"
B <- "number of bootstrap samples to take"
seed_num <- "seed number"
stat <- "statistic that you want to look at" # for now this can be mean, sd, var
bs_samps <- vector("list", B)
set.seed(seed_num)
for(j in 1:B) 
  {bs_samps[[j]] <- sample(X, size = length(X), replace = T)}
bs_reps <- unlist(lapply(bs_samps, stat))
hist(bs_reps)
sd(bs_reps)

## Bootstrapping Examples
####################################################################################
Example 1:
Construct the bootstrap distribution for the sample mean of the weightloss values using the full sample of 100 subjects.  Use B = 1000 bootstrap samples.
####################################################################################
wloss_data <- read.table("/Volumes/Passport/Teaching/P6104 Intro Online/P6104 Online 2016/Weekly Resources/Module 04/Lecture Notes Module 04/Weightloss.txt", header = T)
X <- wloss_data$loss
B <- 1000
seed_num <- 1234
stat <- mean
bs_samps <- vector("list", B)
set.seed(seed_num)
for(j in 1:B) {
	bs_samps[[j]] <- sample(X, size = length(X), replace = T)
}
bs_reps <- unlist(lapply(bs_samps, stat))
hist(bs_reps)
sd(bs_reps)

####################################################################################
Example 2:
Construct the bootstrap distribution for the sample standard deviation of the weightloss values using the full sample of 100 subjects.  Use B = 1000 bootstrap samples.
####################################################################################
wloss_data <- read.table("/Volumes/Passport/Teaching/P6104 Intro Online/P6104 Online 2016/Weekly Resources/Module 04/Lecture Notes Module 04/Weightloss.txt", header = T)
X <- wloss_data$loss
B <- 1000
seed_num <- 1234
stat <- sd
bs_samps <- vector("list", B)
set.seed(seed_num)
for(j in 1:B) {
	bs_samps[[j]] <- sample(X, size = length(X), replace = T)
}
bs_reps <- unlist(lapply(bs_samps, stat))
hist(bs_reps)
sd(bs_reps)
